On the error performance of the $A_n$ lattices

### 2011-11-02 paper.tex
Added bib and other missing bits.
Added Emanuele's changes.

\usepackage{../bib/mathbf-abbrevs}

@@

%opening
\title{Coding with the Lattice $A_n$}
\author{Robby McKilliam, Ramanan Subramanian, Emanuele Viterbo}

@@

\begin{abstract}
Lattices can be used to construct block codes.  The special family of root lattices has attracted particular attention for this purpose.  We consider the root lattice $A_n$ and derive formulae for the moments of its Voronoi cell.  These formulae enable accurate prediction of the performance of codes constructed from $A_n$.  In two dimensions $A_2$ is the hexagonal lattice and in three dimensions $A_3$ is the face-centered cubic lattice.  These are the densest known sphere packings in dimensions two and three.  Our results automatically include codes constructed using these packings.
\end{abstract}

\begin{keywords}
Lattices, lattice decoding, root lattice, probability of error.
\end{keywords}

@@

We consider block codes for the Gaussian channel constructed from a subset of points from a Euclidean lattice using \emph{lattice decoding} (as opposed to maximum likelihood decoding)~\cite{Erex2004_lattice_decoding}.  It is known that such codes can reach the capacity of the Gaussian channel~\cite{Buda1989_some_opt_codes_structure,Erez2005,Erex2004_lattice_decoding}.  These capacity results make use of randomly generated lattices that do not have any obvious exploitable structure for efficient decoding and encoding~\cite{Jalden2005_sphere_decoding_complexity}.  For this reason (and many others) the study of specific lattices with known structure is important.  The \emph{root lattices} $A_n$, $D_n$, $E_6$, $E_7$ and $E_8$ have attracted particular attention for this purpose~\cite{SPLAG}.  In this paper we consider codes constructed from the root lattice $A_n$ and derive formulae for accurately predicting the performance of these codes.  This is achieved by deriving formulae for the \emph{moments} of the \emph{Voronoi cell} of $A_n$.  Conway and Sloane suggested this type of procedure would be possible in~\cite{Conway1982VoronoiRegions}.  This paper can be seen as a continuation of that work focusing particularly on the lattice $A_n$.

In two dimensions $A_2$ is the hexagonal lattice and in three dimensions $A_3$ is the face-centered cubic lattice.  These are the densest known sphere packings in dimensions two and three and our results automatically include low dimensional codes constructed using these packings.  Unfortunately, but not surprisingly, the lattice $A_n$ does not produce good codes in large dimensions.  We are hopeful that the techniques developed here may be applied to other lattices that do produce good codes.  In particular our techniques might apply to the Coxeter lattices $A_n^m$ of which the root lattice $E_8 \simeq A_8^3$ is a member~\cite{Coxeter1951,Martinet2003,McKilliam2009CoxeterLattices}.

This paper is organised as follows.  In Section~\ref{sec:latt-latt-codes} we give a brief overview of lattices and codes constructed from them, i.e. \emph{lattice codes}.  We describe lattice decoding and show how the probability of coding error can be expressed in terms of the moments of the \emph{Voronoi cell} of the lattice used.  Section~\ref{sec:main-result} states the main result without any proof.  We give recursive formula to compute the moments of the Voronoi cell of $A_n$.  Section~\ref{sec:lattice-a_n} describes the lattice $A_n$ and some of its properties.  An important property for our purposes is that the Voronoi cell of $A_n$ is precisely the projection of a $n$-dimensional hypercube orthogonal to one of its vertices~\cite{McKilliam2009CoxeterLattices,McKilliam2010thesis}.  %This result was previously known~\cite{McKilliam2009CoxeterLattices,McKilliam2010thesis}.  The proof is short and we give it again here so that this paper is self contained.  
In Section~\ref{sec:integr-funct-over} we use this property to show how integrals over the Voronoi cell of $A_n$ can be expressed as integrals over the $n$-dimensional hypercube. These integrals are solvable and we use them to obtain the moments of the Voronoi cell in Section~\ref{sec:powers-eucl-norm}.  In  Section~\ref{sec:results-simulations} we plot the probability of error versus signal to noise ratio for codes constructed from the lattices $A_1 \simeq \ints$, $A_2$, $A_3 \simeq D_3$ and $A_4$.  We also plot the results of Monte-Carlo simulations.  These simulations corroborate with our analytical results.  %An advantage of the analytical results is that the probability of error can calculated at high SNR, when the probabilty of error is small.  This is in contrast to Monte-Carlo simulations that are computationally infeasible at when the probability of error is small.  We display the performance of these lattice codes

@@

The Voronoi cell encodes many interesting lattice properties such as the packing radius, covering radius, kissing number, center density, thickness, and the normalized second moment (or quantising constant)~\cite{Viterbo_diamond_cutting_1996, SPLAG}. The error probability of a lattice code can also be evaluated from the Voronoi cell as we will see.  There exist algorithms to completely enumerate the Voronoi cell of an arbitrary lattice~\cite{Viterbo_diamond_cutting_1996,Sikiric_complex_algs_vor_cells_2009,Sikiric_vor_reduction_covering_2008,Valentin2003_coverings_tilings_low_dimension}.  In general these algorithms are only computationally feasible when the dimension is small (approximately $n \leq 9$).  Even with a complete description of the Voronoi cell it is not necessarily easy to compute the probability of coding error.

The Voronoi cell is linked with the problem of \emph{lattice decoding}.  Given some point $\ybf \in \reals^n$ a \emph{lattice decoder} (or \emph{nearest lattice point algorithm}) returns the lattice point in $\Lambda$ that is nearest to $\ybf$~\cite{Agrell2002}.  Equivalently it returns the lattice point $\xbf$ such that the translated Voronoi cell $\vor(\Lambda) + \xbf$ contains $\ybf$.  Computationally lattice decoding is known to be NP-hard under certain conditions when the lattice itself, or rather a basis thereof, is considered as an additional input parameter~\cite{micciancio_hardness_2001, Jalden2005_sphere_decoding_complexity}. Nevertheless, algorithms exist that can compute the nearest lattice point in reasonable time if the dimension is small (approximately $n \leq 60$). One such algorithm is the \term{sphere decoder}~\cite{Viterbo_sphere_decoder_1999,Pohst_sphere_decoder_1981,Agrell2002}.   %Kannan~\cite{Kannan1987_fast_general_np} suggested a different approach that is known to be asymptotically faster than the sphere decoder. A good overview of these techniques is given by~\cite{Agrell2002}.
Fast nearest point algorithms are known for specific lattices~\citep{Conway1982FastQuantDec, McKilliam2008, McKilliam2009CoxeterLattices,Vardy1993_leech_lattice_MLD}. For example, the root lattices $D_n$ and $A_n$ and their dual lattices $D_n^*$ and $A_n^*$ can be decoded in linear-time, i.e. in a number of operations of order $O(n)$~\cite{Conway1982FastQuantDec,McKilliam2009CoxeterLattices}.

@@

%For each lattice point $\ybf \in \Lambda$ there exists a Voronoi cell containing those point from $\reals^n$ nearer to $\ybf$ that any other lattice point.  Due to the structure of the lattice the Voronoi cell of each lattice point is the same shape.  Denote by $\vor(\Lambda)$ the Vornoi cell of the lattice point at the origin. NEAREST POINT PROBLEM.
\newcommand{\calX}{\mathcal X}
Lattices can be used to construct \emph{block codes}.  A block code $\calX$ of length $n$ is a finite set of vectors from $\reals^n$.  Each vector in $\calX$ is called a \emph{codeword}.  The number of codewords is denoted by $\abs{\calX}$.  %If each codeword is transmitted with equal probability then the rate of the code is $R = B\frac{1}{n}\log_2\abs{\calX}$ where $B$ is the symbol rate or bandwidth, i.e. the rate in Hertz at which the elements (or symbols) in a codeword are transmitted.  
The average power of the code is given by $P = \frac{1}{\abs{\calX}}\sum_{\xbf \in \calX}\|\xbf\|^2$.  In the additive white Gaussian channel the received signal takes the form
\[
\ybf = \xbf_0 + \wbf
\]
where $\ybf \in \reals^n$, $\xbf_0 \in \calX$ and $\wbf$ is a vector of independent and identically distributed Gaussian random variables with variance $\sigma^2$.  If the receiver employs maximum likelihood decoding then the estimator of $\xbf$ given $\ybf$ at the receiver is
\begin{equation}\label{eq:mldecoder}
\hat{\xbf} = \min_{\xbf \in \calX} \| \ybf - \xbf \|^2.
\end{equation}
That is, the receiver computes the codeword in $\calX$ nearest in Euclidean distance to the received signal $\ybf$.  Assuming that each codeword is transmitted with equal probability then the probability of correct decoding is
\[
\sum_{\xbf \in \calX} \frac{\prob( \hat{\xbf} = \xbf )}{\abs{\calX}}.
\]
The probability of error is $1 - \frac{1}{\abs{\calX}} \sum_{\xbf \in \calX} \prob( \hat{\xbf} = \xbf )$.  %The maximum possible rate of reliable communications in this channel is given by the channel capacity $B\log_2\left(1 + \frac{P}{\sigma^2}\right)$.  If the codewords are chosen uniformly randomly on a sphere of radius $\sqrt{P}$ then as $n$ gets large the rate of the code can be made arbitrarily close to the channel capacity while simultaneously the probability of error can be made arbitrarily small.  Another way of stating this is that we can choose close to  $2^{nR/B}$ codewords uniformly randomly on a sphere of radius $\sqrt{P}$ and as $n$ gets large the probability of error decreases to zero.

@@

A \emph{lattice code} is a block code with codewords given by a finite subset of points from some lattice $\Lambda$ in $\reals^n$.  There are an infinity of ways to choose a finite subset from a lattice, but common approaches make use of a bounded subset of $\reals^n$ called a \emph{shaping region}.  The lattice points in $\Lambda$ that intersect the shaping region make up the code.  That is, if $S \subset \reals^n$ is a shaping region then our codewords are given by
\[
\calX = S \cap \Lambda.
\]
Possible choices of shaping region are $n$-dimensional spheres, spherical shells, hypercubes or the Voronoi cell of a \emph{sublattice} of $\Lambda$~\cite{Buda1989_some_opt_codes_structure,Erex2004_lattice_decoding,Conway1983VoronoiCodes}. A consequence of the finiteness of a lattice code is that the maximum likelihood decoder~\eqref{eq:mldecoder} is \emph{not} equivalent to lattice decoding.  Computing a nearest lattice point will not in general return a lattice point from the code (the decoded lattice point might lie outside the shaping region).  A question that remained open for some time was whether lattice codes employing lattice decoding could achieve channel capacity.  This question was answered in the affirmative by Erez and Zamir~\cite{Erex2004_lattice_decoding}.

Let $\calX$ be a lattice code constructed from a lattice $\Lambda$.  Assuming that each codeword $\xbf \in \calX$ has an equal probability of being transmitted and that the receiver employs lattice decoding, then the average probability of correct decoding in the AWGN channel is\footnote{In order to approach capacity Erez and Zamir~\cite{Erex2004_lattice_decoding} employ a uniform dither and also a minimum mean square error scaling parameter $\alpha$ (the so called \emph{modulo-lattice additive noise channel}).  A result of this is that that observed noise distribution is no longer Gaussian.  However, for large signal to noise ratio $\alpha \approx 1$ and the observed noise is very close to Gaussian.  In this case~\eqref{eq:PEgaussian} is a reasonable approximation.}
\begin{align}
P_C &= \frac{1}{\abs{\calX}} \sum_{\cbf\in \calX} \prob( \xbf + \wbf \in \vor(\Lambda) + \xbf )  \\
&= \prob(\xbf \in \vor(\Lambda) ) \nonumber \\
&=   \frac{1}{(\sqrt{2\pi}\sigma)^n} \int_{\vor(\Lambda)} e^{-\|\xbf\|^2 / 2\sigma^2 } d\xbf. \label{eq:PEgaussian}
\end{align}
The probability of error is $P_E = 1 - P_C$.  By expanding $e^x = 1  + x + \frac{x^2}{2} + \dots$ we obtain
\begin{align}
P_C &= \frac{1}{(\sqrt{2\pi}\sigma)^n}\int_{\vor(\Lambda)} 1 - \frac{\|\xbf\|^2}{2\sigma^2} + \frac{\left(\|\xbf\|^2\right)^2}{ 4\sigma^42!} - \dots d\xbf \nonumber \\
&= \frac{1}{(\sqrt{2\pi}\sigma)^n} \sum_{m=0}^\infty \frac{(-1)^m}{2^m\sigma^{2m}m!} \int_{\vor(\Lambda)} \left(\|\xbf\|^2\right)^m d\xbf.  \label{eq:summomentproberror}
\end{align}
So, to obtain arbitrarily accurate approximations to the probability of error it is enough to know the values of $\int_{\vor(\Lambda)} \left(\|\xbf\|^2\right)^m d\xbf$ for $m=1,2\dots,M$ for some sufficiently large $M$.

In this paper we focus on lattice codes constructed from the family of lattices called $A_n$. We will discover expressions for
\[
 C_n(m) = \int_{\vor(A_n)} \left(\|\xbf\|^2\right)^m d\xbf.
\]
These can be summed to give arbitrarily accurate approximations for the probability of error.  In this paper we call $C_n(m)$ the $m$th \emph{moment} of the lattice $A_n$.

@@

We now state our main result.  The $m$th moment $C_n(m)$ satisfies
\begin{equation}\label{eq:theCmformula}
\frac{C_n(m)}{m!} = \frac{n\sqrt{n+1}}{n+2m}\sum_{k=0}^{m}\sum_{a =0}^{k}\sum_{b=0}^{k-a} \frac{G(n-1,a,2k - 2a - b)}{(-1)^{k-a}H(n,m,k,a,b)} 
\end{equation}
where the function
\[
H(n,m,k,a,b) = \frac{(n+1)^{m-a}a!(m-k)!b! (k-a-b)!}{2^{b} n^{m-k}},
\]
and the function $G(n,c,d)$ satisfies the recursion
\begin{equation}\label{eq:theGrecursion}
G(n,c,d) = \sum_{c'=0}^{c} \sum_{d'=0}^{d} \binom{c}{c'}\binom{d}{d'} \frac{G(n-1,c-c',d-d')}{2c'+d'+1}
\end{equation}
with the initial conditions
\[
G(1,c,d) = \frac{1}{2c+d+1} \qquad \text{and} \qquad G(n,0,0) = 1.
\]
For fixed $m$ it is possible to solve this recursion in $n$ and obtain formula for the $C_m(n)$ in terms of $n$.  The first four such formula are:
\begin{align*}
C_0(n) &= \sqrt{n+1} \qquad \text{the volume of } \vor(A_n),\\
C_1(n) &= \frac{n(n+3)}{12\sqrt{n+1}} \qquad \text{the second moment of $\vor{A_n}$~ \cite[p. 462]{SPLAG}}, \\
C_2(n) &=  \frac{50 n + 55 n^2 + 34 n^3 + 5 n^4}{720 (1 + n)^{3/2}}, \\
C_3(n) &= \frac{1960 n + 2142 n^2 + 2681 n^3 + 1423 n^4 + 399 n^5 + 35 n^6}{60480 (1 + n)^{5/2}}.
%C_4(n) &= \frac{93744 n + 34356 n^2 + 112172 n^3 + 89343 n^4 + 53224 n^5 + 17246 n^6 + 2940 n^7 + 175 n^8}{3628800 (1 + n)^{7/2}}.
\end{align*}
We have tabulated these formula for $m=0$ to $40$.  We derive these results in Section~\ref{sec:powers-eucl-norm}.  First we need some properties of the lattice $A_n$.

@@

\small
\bibliography{../bib/bib}

>>>>

\usepackage{mathbf-abbrevs}

@@

%opening
\title{On the error performance of the $A_n$ lattices}
\author{Robby McKilliam, Ramanan Subramanian, Emanuele Viterbo}

@@

\begin{abstract}
We consider the root lattice $A_n$ and derive explicit formulae for
the moments of its Voronoi cell.  We then show that these formulae
enable accurate prediction of the error probability performance of
lattice codes constructed from $A_n$.
\end{abstract}

\begin{keywords}
Lattices, lattice decoding, root lattice, probability of error, Voronoi cell.
\end{keywords}

@@

The \emph{root lattices} $A_n$, $D_n$, $E_6$, $E_7$ and $E_8$ have
attracted particular attention as structured codes for the AWGN
channel~\cite{SPLAG}. The highly symmetric structure of these lattices
provides the grounds for extremely efficient decoding algorithms.  In
this paper we consider codes constructed from the root lattice $A_n$
and derive formulae for accurately predicting the performance of these
codes.  This is achieved by deriving formulae for the \emph{moments}
of the \emph{Voronoi cell} of $A_n$.  Conway and Sloane suggested this
approach to compute the quantizing constants (second order moments) of
the root lattices,~\cite{Conway1982VoronoiRegions}.  In this paper we
extend their technique to compute the moments of any order for lattice
$A_n$.

In two dimensions $A_2$ is the hexagonal lattice and in three
dimensions $A_3$ is the face-centered cubic lattice.  These are the
densest known sphere packings in dimensions two and three and our
results automatically include low dimensional codes constructed using
these packings.  In general, the lattice $A_n$ does not produce
asymptotically good codes in large dimensions, but still offer a
significant coding gain in small dimensions.  For these cases, we
provide a error probability expression which can be computed to any
degree of accuracy at any finite signal-to-noise ratio.
%We are hopeful that the techniques developed here may be applied to other lattices that do produce good codes.  In particular our techniques might apply to the Coxeter lattices $A_n^m$ of which the root lattice $E_8 \simeq A_8^3$ is a member~\cite{Coxeter1951,Martinet2003,McKilliam2009CoxeterLattices}.

This paper is organised as follows.  In Section~\ref{sec:latt-latt-codes} we give a brief overview of lattices
and codes constructed from them, i.e., \emph{lattice codes}.  We
describe lattice decoding and show how the probability of coding error
can be expressed in terms of the moments of the \emph{Voronoi cell} of
the lattice used.  Section~\ref{sec:main-result} states the main
result (without proof), which gives recursive formula to compute the
moments of the Voronoi cell of $A_n$.  Section~\ref{sec:lattice-a_n}
describes the lattice $A_n$ and some of its properties.  An important
property for our purposes is that the Voronoi cell of $A_n$ is
precisely the projection of a $n+1$-dimensional hypercube orthogonal
to one of its vertices~\cite{McKilliam2009CoxeterLattices,McKilliam2010thesis}.
 %This result was previously known~\cite{McKilliam2009CoxeterLattices,McKilliam2010thesis}.  The proof is short and we give it again here so that this paper is self contained.
In Section~\ref{sec:integr-funct-over} we use this property to show
how integrals over the Voronoi cell of $A_n$ can be expressed as
integrals over the $n$-dimensional hypercube. These integrals are
solvable and we use them to obtain the moments of the Voronoi cell in
Section~\ref{sec:powers-eucl-norm}.  In
Section~\ref{sec:results-simulations} we plot the probability of error
versus signal to noise ratio for codes constructed from the lattices
$A_1 \simeq \ints$, $A_2$, $A_3 \simeq D_3$ and $A_4$.  We also plot
the results of Monte-Carlo to  corroborate our analytical results.
%An advantage of the analytical results is that the probability of error can calculated at high SNR, when the probabilty of error is small.  This is in contrast to Monte-Carlo simulations that are computationally infeasible at when the probability of error is small.  We display the performance of these lattice codes

@@

The Voronoi cell encodes many interesting lattice properties such as
the packing radius, covering radius, kissing number, center density,
thickness, and the normalized second moment (or quantising
constant)~\cite{Viterbo_diamond_cutting_1996, SPLAG}. The error
probability of a lattice code can also be evaluated from the Voronoi
cell as we will see.  There exist algorithms to completely describe
the Voronoi cell of an arbitrary
lattice~\cite{Viterbo_diamond_cutting_1996,Sikiric_complex_algs_vor_cells_2009,Sikiric_vor_reduction_covering_2008,Valentin2003_coverings_tilings_low_dimension}.
 In general these algorithms are only computationally feasible when
the dimension is small (approximately $n \leq 9$).  Even with a
complete description of the Voronoi cell it is not generally easy to
compute the probability of error of a lattice in an additive white
Gaussian noise (AWGN) channel.

The Voronoi cell is linked with the problem of \emph{lattice
decoding}.  Given some point $\ybf \in \reals^n$ a \emph{lattice
decoder} (or \emph{nearest lattice point algorithm}) returns the
lattice point in $\Lambda$ that is nearest to
$\ybf$~\cite{Agrell2002}.  Equivalently it returns the lattice point
$\xbf$ such that the translated Voronoi cell $\vor(\Lambda) + \xbf$
contains $\ybf$.  Computationally lattice decoding is known to be
NP-hard under certain conditions when the lattice itself, or rather a
basis thereof, is considered as an additional input
parameter~\cite{micciancio_hardness_2001,
Jalden2005_sphere_decoding_complexity}. Nevertheless, algorithms exist
that can compute the nearest lattice point in reasonable time if the
dimension is small (approximately $n \leq 60$). One such algorithm is
the \term{sphere decoder}~\cite{Viterbo_sphere_decoder_1999,Pohst_sphere_decoder_1981,Agrell2002}.
 %Kannan~\cite{Kannan1987_fast_general_np} suggested a different approach that is known to be asymptotically faster than the sphere decoder. 
A good overview of these techniques is given by~\cite{Agrell2002}.
Fast nearest point algorithms are known for specific
lattices~\cite{Conway1982FastQuantDec, McKilliam2008,
McKilliam2009CoxeterLattices,Vardy1993_leech_lattice_MLD}. For
example, the root lattices $D_n$ and $A_n$ and their dual lattices
$D_n^*$ and $A_n^*$ can be decoded in linear-time, i.e. in a number of
operations of order $O(n)$~\cite{Conway1982FastQuantDec,McKilliam2009CoxeterLattices}.

@@

%For each lattice point $\ybf \in \Lambda$ there exists a Voronoi cell containing those point from $\reals^n$ nearer to $\ybf$ that any other lattice point.  Due to the structure of the lattice the Voronoi cell of each lattice point is the same shape.  Denote by $\vor(\Lambda)$ the Vornoi cell of the lattice point at the origin. NEAREST POINT PROBLEM.
\newcommand{\calX}{\mathcal X}
Lattices can be used to construct \emph{signal space codes}.  A signal
space code $\calX$ of dimension $n$ is a finite set of vectors (or
points) in the Euclidean space $\reals^n$.  Each vector in $\calX$ is
called a \emph{codeword} and represents a particular signal.  The
number of codewords is denoted by $\abs{\calX}$.  If each codeword is
transmitted with equal probability then the rate of the code is $R =
\frac{1}{n}\log_2\abs{\calX}$ bits per codeword.
The average power of the code is given by $P =
\frac{1}{\abs{\calX}}\sum_{\xbf \in \calX}\|\xbf\|^2$.  In the AWGN
channel the received signal takes the form
\[
\ybf = \xbf + \wbf
\]
where $\ybf \in \reals^n$, $\xbf \in \calX$ and $\wbf$ is a vector of
independent and identically distributed Gaussian random variables with
variance $\sigma^2$.  If the receiver employs maximum likelihood
decoding then the estimator of $\xbf$ given $\ybf$ at the receiver is
\begin{equation}\label{eq:mldecoder}
\hat{\xbf} = \mbox{arg~}\min_{\xbf \in \calX} \| \ybf - \xbf \|^2.
\end{equation}
That is, the receiver computes the codeword in $\calX$ nearest in
Euclidean distance to the received signal $\ybf$.  Assuming that each
codeword is transmitted with equal probability then the probability of
correct decoding is
\[
P_C =\frac{1}{\abs{\calX}}\sum_{\xbf \in \calX} \prob( \hat{\xbf} = \xbf ).
\]
and the probability of error is
\[
P_E = 1 - \frac{1}{\abs{\calX}} \sum_{\xbf \in \calX} \prob(
\hat{\xbf} = \xbf ).
\]
%The maximum possible rate of reliable communications in this channel is given by the channel capacity $B\log_2\left(1 + \frac{P}{\sigma^2}\right)$.  If the codewords are chosen uniformly randomly on a sphere of radius $\sqrt{P}$ then as $n$ gets large the rate of the code can be made arbitrarily close to the channel capacity while simultaneously the probability of error can be made arbitrarily small.  Another way of stating this is that we can choose close to $2^{nR/B}$ codewords uniformly randomly on a sphere of radius $\sqrt{P}$ and as $n$ gets large the probability of error decreases to zero.

@@

A \emph{lattice code} is a signal space code with codewords taken from
a finite subset of points of some lattice $\Lambda$ in $\reals^n$.
There are infinitely many ways to choose a finite subset from a
lattice, but common approaches make use of a bounded subset of
$\reals^n$, called a \emph{shaping region} $S \subset \reals^n$ and
\[
\calX = S \cap \Lambda.
\]
Possible choices of shaping region are $n$-dimensional spheres,
spherical shells, hypercubes or the Voronoi cell of a
\emph{sublattice} of
$\Lambda$~\cite{Buda1989_some_opt_codes_structure,Erex2004_lattice_decoding,Conway1983VoronoiCodes}.
A consequence of the finiteness of a lattice code is that the maximum
likelihood decoder~\eqref{eq:mldecoder} is \emph{not} equivalent to
lattice decoding.  Computing a nearest lattice point will not in
general return a lattice point from the code (the decoded lattice
point might lie outside the shaping region).
 %A question that remained open for some time was whether lattice codes employing lattice decoding could achieve channel capacity.  This question was answered in the affirmative by Erez and Zamir~\cite{Erex2004_lattice_decoding}.

Let $\calX$ be a lattice code constructed from a lattice $\Lambda$.
Assuming that each codeword $\xbf \in \calX$ has an equal probability
of being transmitted and that the receiver employs lattice decoding,
then the average probability of correct decoding in the AWGN channel
is
%\footnote{In order to approach capacity Erez and Zamir~\cite{Erex2004_lattice_decoding} employ a uniform dither and also a minimum mean square error scaling parameter $\alpha$ (the so called \emph{modulo-lattice additive noise channel}).  A result of this is that that observed noise distribution is no longer Gaussian. However, for large signal to noise ratio $\alpha \approx 1$ and the observed noise is very close to Gaussian.  In this case~\eqref{eq:PEgaussian} is a reasonable approximation.}
\begin{align}
P_C &\leq P_C(\Lambda) = \lim_{\abs{\calX}\rightarrow \infty}
 \frac{1}{\abs{\calX}} \sum_{\xbf\in \calX} \prob( \xbf + \wbf \in
\vor(\Lambda) + \xbf )  \\
&= \prob(\xbf \in \vor(\Lambda) ) \nonumber \\
&=   \frac{1}{(\sqrt{2\pi}\sigma)^n} \int_{\vor(\Lambda)}
e^{-\|\xbf\|^2 / 2\sigma^2 } d\xbf. \label{eq:PEgaussian}
\end{align}
The probability of error is $P_E \leq 1 - P_C(\Lambda)$, where the
upper bound is asymptotically tight for large constellations (the
contribution of the signal set points on near the bounding region
boarder can be assumed negligible).  By expanding $e^x = 1  + x +
\frac{x^2}{2} + \dots$ we obtain
\begin{align}
P_C(\Lambda)  &= \frac{1}{(\sqrt{2\pi}\sigma)^n}\int_{\vor(\Lambda)} 1
- \frac{\|\xbf\|^2}{2\sigma^2} + \frac{\left(\|\xbf\|^2\right)^2}{
4\sigma^42!} - \dots d\xbf \nonumber \\
&= \frac{1}{(\sqrt{2\pi}\sigma)^n} \sum_{m=0}^\infty
\frac{(-1)^m}{2^m\sigma^{2m}m!} \int_{\vor(\Lambda)} \|\xbf\|^{2m}
d\xbf.  \label{eq:summomentproberror}
\end{align}
So, to obtain arbitrarily accurate approximations to the probability
of error it is enough to know the values of $\int_{\vor(\Lambda)}
\|\xbf\|^{2m} d\xbf$ for $m=1,2\dots$ for some sufficiently large $m$.
The number of terms required will increase as the noise variance gets
smaller. This implies that the bound with a fixed number of terms is
not asymptotically tight but is very accurate up to a finite
signal-to-noise ratio.

In this paper we focus on $n$-dimensional lattice codes constructed from the family of lattices called $A_n$ and  we will derive expressions for the $m$th \emph{moment}s
\[
 C_n(m) = \int_{\vor(A_n)} \|\xbf\|^{2m} d\xbf.
\]
These can be summed in (\ref{eq:summomentproberror}) to give arbitrarily accurate approximations for the probability of error.

@@

We now state our main result.  The $m$th moment $C_n(m)$ satisfies
\begin{equation}\label{eq:theCmformula}
\frac{C_n(m)}{m!} = \frac{n\sqrt{n+1}}{n+2m}\sum_{k=0}^{m}\sum_{a =0}^{k}\sum_{b=0}^{k-a} \frac{G(n-1,a,2k - 2a - b)}{(-1)^{k-a}H(n,m,k,a,b)} 
\end{equation}
where the function
\[
H(n,m,k,a,b) = \frac{(n+1)^{m-a}a!(m-k)!b! (k-a-b)!}{2^{b} n^{m-k}},
\]
and the function $G(n,c,d)$ satisfies the recursion
\begin{equation}\label{eq:theGrecursion}
G(n,c,d) = \sum_{c'=0}^{c} \sum_{d'=0}^{d} \binom{c}{c'}\binom{d}{d'} \frac{G(n-1,c-c',d-d')}{2c'+d'+1}
\end{equation}
with the initial conditions
\[
G(1,c,d) = \frac{1}{2c+d+1} \qquad \text{and} \qquad G(n,0,0) = 1.
\]
For fixed $m$ it is possible to solve this recursion in $n$ and obtain formula for the $C_m(n)$ in terms of $n$.  The first four such formula are:
\begin{align*}
C_0(n) &= \sqrt{n+1} \qquad \text{the volume of } \vor(A_n),\\
C_1(n) &= \frac{n(n+3)}{12\sqrt{n+1}} \qquad \text{the second moment of $\vor{A_n}$~ \cite[p. 462]{SPLAG}}, \\
C_2(n) &=  \frac{50 n + 55 n^2 + 34 n^3 + 5 n^4}{720 (1 + n)^{3/2}}, \\
C_3(n) &= \frac{1960 n + 2142 n^2 + 2681 n^3 + 1423 n^4 + 399 n^5 + 35 n^6}{60480 (1 + n)^{5/2}}.
%C_4(n) &= \frac{93744 n + 34356 n^2 + 112172 n^3 + 89343 n^4 + 53224 n^5 + 17246 n^6 + 2940 n^7 + 175 n^8}{3628800 (1 + n)^{7/2}}.
\end{align*}
We have explicitly tabulated these formula for $m=0$ to $40$. For larger $m$s direct evaluation for specific $n$ from the recursive formula is preferable.  We will derive these results in Section~\ref{sec:powers-eucl-norm}, but we first need some properties of the lattice $A_n$.

@@

\small
\bibliography{bib}
